A seamless, client-centric programming model for type safe web applications

### 2014-05-07 paper.tex
GWT, Opa, Conductance, Sunroof fleshed out a little.

\paragraph{Conductance} Conductance\ \cite{conductance} is an application
server built on StratifiedJS, a JavaScript language extension which adds a few
niceties such as cooperative multitasking and more concise syntax for many
common tasks.
Conductance uses an RPC-based model for client-server communication, much like
our own, but also adds the possibility for the server to independently transmit
data back to the client through the use of shared variables or call back into
the client by way of function objects received via RPC call, as well as the
possibility for both client and server to seamlessly modify variables located
on the opposite end of the network. Conductance is quite new and has no
relevant publications. It is, however, used for several large scale web
applications.

@@

\paragraph{Opa} The Opa framework\ \cite{opa}, much like Conductance, enables
seamless communication within the JavaScript language itself.
While Opa improves upon Conductance by introducing non-mandatory type checking
to the JavaScript world, its communication model is based on implicit
information flows, allowing the server to read and update mutable state on the
client and vice versa. While this is a quite flexible programming model, we
believe that this uncontrolled, implicit information flow makes programs harder
to follow, debug, secure and optimize.

\paragraph{Google Web Toolkit} Google Web Toolkit\ \cite{gwt}, a Java
compiler targeting the browser, provides its own solution to client-server
interoperability as well. This solution is based on callbacks, forcing
developers to code in a confusing continuation passing style. It also suffers
from excessive boilerplate code and an error-prone configuration process.

@@

Sunroof is completely type-safe, in the DSL itself as well as in the
communication with the Haskell host. However, the fact that client and server
must be written in two separate languages - any code used to generate
JavaScript must be built solely from the primitives of the Sunroof
language in order to be compilable into JavaScript, precluding use of general
Haskell code - makes code reuse hard.

@@

\paragraph{Advantages to our approach} We believe that our approach has a
number of distinct advantages to the aforementioned attacks on the problem.

@@

\paragraph{Client-centricity} Unlike Conductance and Duetto, our approach
takes a firm stand, regarding the client as the driver in the client-server
relationship with the server taking on the role of a passive computational
or storage resource. The server may thus not call back into the client at
arbitrary points but is instead limited to returning answers to client side
queries. This is clearly less flexible than the back-and-forth model of Sunroof
and Duetto or the shared variables of Conductance. However, we believe that
this restriction makes program flow easier to follow and comprehend. Like the
immutability of Haskell, this model gives programmers a not-so-subtle hint
as to how they may want to structure their programs. Extending our existing
model with an \lstinline!onClient! counterpart to \lstinline!onServer! would be
a simple task, but we are not quite convinced that there is value in doing so.

>>>>

\paragraph{Conductance and Opa} Conductance\ \cite{conductance} is an application
server built on StratifiedJS, a JavaScript language extension which adds a few
niceties such as cooperative multitasking and more concise syntax for many
common tasks.
Conductance uses an RPC-based model for client-server communication, much like
our own, but also adds the possibility for the server to independently transmit
data back to the client through the use of shared variables or call back into
the client by way of function objects received via RPC call, as well as the
possibility for both client and server to seamlessly modify variables located
on the opposite end of the network. Conductance is quite new and has no
relevant publications. It is, however, used for several large scale web
applications.

@@

The Opa framework\ \cite{opa}, another JavaScript framework, is an improvement
over Conductance by introducing non-mandatory type checking to the JavaScript
world. Its communication model is based on implicit information flows, allowing
the server to read and update mutable state on the client and vice versa.
While this is a quite flexible programming model, we believe that this
uncontrolled, implicit information flow makes programs harder to follow, debug,
secure and optimize.

\paragraph{Google Web Toolkit} Google Web Toolkit\ \cite{gwt}, a Java
compiler targeting the browser, provides its own solution to client-server
interoperability as well. This solution is based on callbacks, forcing
developers to write code in a continuation passing style. It also suffers
from excessive boilerplate code and an error prone configuration process.
The programming model shares Haste.App's client centricity, relegating the
server to serving client requests.

@@

Sunroof is completely type-safe, in the DSL itself as well as in the
communication with the Haskell host. However, the fact that client and server
must be written in two separate languages - any code used to generate
JavaScript must be built solely from the primitives of the Sunroof
language in order to be compilable into JavaScript, precluding use of general
Haskell code - makes code reuse hard. As the JavaScript DSL is executed from
a native Haskell host, Sunroof's programming model can be said to be somewhat
server centric, but with quite some flexibility due to its back and forth
communication model.

@@

\paragraph{Advantages of our approach} We believe that our approach has a
number of distinct advantages to the aforementioned attacks on the problem.

@@

\paragraph{Client-centricity} Unlike most related work, our approach
takes a firm stand, regarding the client as the driver in the client-server
relationship with the server taking on the role of a passive computational
or storage resource. The server may thus not call back into the client at
arbitrary points but is instead limited to returning answers to client side
queries. This is clearly less flexible than the back-and-forth model of Sunroof
and Duetto or the shared variables of Conductance. However, we believe that
this restriction makes program flow easier to follow and comprehend. Like the
immutability of Haskell, this model gives programmers a not-so-subtle hint
as to how they may want to structure their programs. Extending our existing
model with an \lstinline!onClient! counterpart to \lstinline!onServer! would be
a simple task, but we are not quite convinced that there is value in doing so.
